Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-15 arxiv/diff_cov.tex
polish

\paragraph{Bias-variance equations.}
Our key result characterizes the asymptotic limit of the inverse of the sum of two arbitrarily different sample covariance matrices.
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By Proposition 1 of \citet{WZR20}, we consider the case of $r = 1 < t = 2$, since when $r > 1$, the global minimum of $f(A, B)$ is equivalent to single-task learning.
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
To motivate our study, we consider a special case where $A_1=A_2=1$.
Hence the HPS estimator is equal to $B$.
%Hence we can write down a closed form equation for any local minimizer of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the estimator for task two as follows:
\begin{align}
	\hat{\beta}_2^{\MTL} &= {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \ \ \text{ where } \nonumber \\
	\hat{\Sigma} &= {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariance matrices, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

@@

\begin{theorem}\label{thm_main_RMT}
	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have that
	the variance equation \eqref{eq_variance_2task} satisfies the following estimate:
			\begin{align}\label{lem_cov_shift_eq}
				\resizebox{0.869\hsize}{!}{%
				$\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{ {\hat{\Sigma}^{-1}} - \frac{(a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}{n_1+n_2} }}}
				\le p^{-c_{\varphi}}$
				}%
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				&a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\label{eq_a12extra000}\\
				&a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
\end{theorem}
Due to space limit, we defer the bias limit result to Appendix \eqref{appendix RMT}.
Our result extends Fact \ref{fact_tr} to the inverse of the sum of two sample covariance matrices.
To see this, when $n_1$ is zero, we solve equations \eqref{eq_a12extra000} and  \eqref{eq_a12extra} to obtain that $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply them to equation \eqref{lem_cov_shift_eq}.
For general $A_1,A_2$ that are not equal to one, we can still apply our result by rescaling $X^{(1)}$ and $M$ with $A_1 / A_2$.
We defer a proof sketch of Theorem \ref{thm_main_RMT} until the end of the section.
%This amounts to replacing $M$ with $\frac{A_1}{A_2}M$ in Theorem \ref{thm_main_RMT}.

@@

In the above equation, the $d^2$ scaling term is the bias limit and the $\sigma^2$ scaling term is the variance limit.
This result allows for a more concrete interpretation since the dependence on datasets' properties is explicit.
The proof of Corollary \ref{cor_MTL_loss} can be found in Appendix \ref{app_iso_cov}.
As a remark, %in equation \eqref{cor_MTL_error}, the predication loss $L(\hat{\beta}_2^{\MTL}) $ was obtained using the global minimizer $(\hat A,\hat B)$. B
by combining the bias and variance limits, we can also obtain a bias-variance tradeoff for any local minimizer of $f(A, B)$.
The proof is similar to Corollary \ref{cor_MTL_loss}, so we omit the details.

@@

\begin{example}[Covariate shift]\label{ex_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example focuses on how varying covariate shifts impacts the \textit{variance} limit in equation \eqref{lem_cov_shift_eq}, which gives that
%in the left hand side of equation \eqref{lem_cov_shift_eq}:
\begin{align*}
	\bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} &\rightarrow \frac{1}{n_1 + n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}} \\
	&= \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}.
\end{align*}
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Hence the variance limit is determined by the spectrum of $M$. % To be clear, for this example we assume that the bias is $0$.
To illustrate the above result, suppose that half of $M$'s singular values are equal to $\lambda > 1$ and the other half are equal to $\lambda^{-1}$.
In particular, when $\lambda = 1$, there is no covariate shift.
As $\lambda$ increases, the severity of covariate shift increases.
We observe the following dichotomy.
%\begin{itemize}

@@

We explain why the dichotomy happens. The variance limit in this example is equal to
$\frac{p}{2(n_1 + n_2)} f(\lambda)$, where
\[ f(\lambda) = {(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}. \]
Using the fact that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2}$, we can verify
%\begin{align*}
%	f(\lambda) - f(1) &= \left(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&+ \left(\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&- \frac{2(n_1 + n_2)}{n_1 + n_2 - p} \\
%	&= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
%\end{align*}
\begin{align*}
	f(\lambda) - f(1) &= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
\end{align*}
where $g(\lambda, a_1) \ge 0$ is a fixed function. 
%and can be derived from algebraic calculations (details omitted).
We claim that $a_1 \ge \frac{n_1 + n_2-p}{2(n_1 + n_2 )}$ if and only if $n_1 \ge n_2$, which explains the dichonomy. In fact, if $a_1>a_2$, then equation \eqref{eq_a12extra000} gives that $a_1> \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$, and equation \eqref{eq_a12extra} gives that
\begin{align*}
 \frac{n_1}{n_1 + n_2} &> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) >\frac{1}{2}.
\end{align*}
%\begin{align*}
% \frac{n_1}{n_1 + n_2} &=a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} \\
%	&> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) =\frac{1}{2}.
%\end{align*}
This implies $n_1>n_2$. The other direction follows from similar arguments. % Similarly, if $a_1<a_2$, equations  \eqref{eq_a12extra000} and  \eqref{eq_a12extra} give that $a_1 < \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$ and $n_1<n_2$.
%Thus, we conclude that $f(\lambda) \ge f(1)$ if and only if $n_1 \ge n_2$.

@@

\textbf{Overview of Stieltjes transform.}
For the rest of this section, we present an overview of the proof of Theorem \ref{thm_main_RMT}.
The central quantity of interest is the inverse of the sum of two sample covariance matrices.
We note that the variance equation $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}]$ is equal to $(n_1 + n_2)^{-1} \bigtr{W^{-1}}$, where $W$ is
\begin{align}\label{eigen2extra}
	\frac{\Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V}{n_1 + n_2}.
\end{align}
Here $U\Lambda V^\top$ is defined as the SVD of $M$.
This formulation is helpful because we know that $(Z^{(1)})^{\top} Z^{(1)}$ and $(Z^{(2)})^{\top} Z^{(2)}$ are both sample covariance matrices with isotropic population covariance, and $U, V$ are both orthonormal matrices.
For example, if $Z^{(1)},Z^{(2)}$ are both Gaussian random matrices, by rotational invariance, $Z^{(1)} U, Z^{(2)}V$ are still Gaussian random matrices.

@@

\begin{figure*}[!t]
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/same_covariates.eps}
		\caption{Example \ref{ex_same_cov}}
		\label{fig_same_cov}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/sample_ratio_several_d.eps}
		\caption{Example \ref{ex_sample_ratio}}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/covariate_shift.eps}
		\caption{Example \ref{ex_covshift}}
		\label{fig_covariate}
	\end{subfigure}
	\caption{%Three takeaways of our theory in Section \ref{sec_insight}.
	Our estimated losses (solid line) match the empirical losses (dotted line) accurately under various settings in dimension $p = 200$.
	\textbf{Left.} Validating Example \ref{ex_same_cov} for ten tasks: the noise variance $\sigma^2$ is $1/4$.
	\textbf{Middle.} Validating Example \ref{ex_sample_ratio} for two tasks: we discover several interesting phenomena by fixing task two's sample size and increasing task one's sample size.
	Depending on how large $d^2$ is, task two's prediction loss decreases initially before increasing again, or decreases monotonically.
	\textbf{Right.} We show how different levels of covariate shift affect hard parameter sharing when there is no bias.
	Having covariate shift increases task two's prediction loss when task two's sample size is smaller than task one. Otherwise, having covariate shift (surprisingly) decreases task two's prediction loss.}
	\label{fig_model_shift_phasetrans}
\end{figure*}

>>>>

\paragraph{Bias-variance equations.}
Our key result characterizes the asymptotic limit of the inverse of the sum of two arbitrarily different sample covariance matrices.
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
We consider the case of $r = 1 < t = 2$, since when $r > 1$, the global minimum of $f(A, B)$ reduces to single-task learning (cf. Proposition 1 of \citet{WZR20}).
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
To motivate our study, we consider a special case where $A_1=A_2=1$.
Hence the HPS estimator is equal to $B$.
%Hence we can write down a closed form equation for any local minimizer of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the estimator for task two as follows:
\begin{align}
	\hat{\beta}_2^{\MTL} &= {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \ \ \text{ where } \nonumber \\
	\hat{\Sigma} &= {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariance matrices, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

@@

\begin{theorem}\label{thm_main_RMT}
	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have that
	the variance equation \eqref{eq_variance_2task} $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}]$ (leaving out $\sigma^2$) satisfies the following estimate:
			\begin{align}\label{lem_cov_shift_eq}
				\resizebox{0.869\hsize}{!}{%
				$\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{ {\hat{\Sigma}^{-1}} - \frac{(a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}{n_1+n_2} }}}
				\le p^{-c_{\varphi}}$
				}%
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				&a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\label{eq_a12extra000}\\
				&a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
\end{theorem}
Due to space limit, we defer the bias limit result to Appendix \eqref{appendix RMT}.
Our result extends Fact \ref{fact_tr} to the inverse of the sum of two sample covariance matrices.
To see this, when $n_1$ is zero, we solve equations \eqref{eq_a12extra000} and  \eqref{eq_a12extra} to obtain that $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply them to equation \eqref{lem_cov_shift_eq}.
For general $A_1,A_2$ that are not equal to one, we can still apply our result by rescaling $X^{(1)}$ and $M$ with $A_1 / A_2$.
We defer a proof sketch of Theorem \ref{thm_main_RMT} until the end of the section.
%This amounts to replacing $M$ with $\frac{A_1}{A_2}M$ in Theorem \ref{thm_main_RMT}.

@@

In the above inequality, the $d^2$ scaling term is the bias limit and the $\sigma^2$ scaling term is the variance limit.
This result allows for a more concrete interpretation since the dependence on datasets' properties is explicit.
The proof of Corollary \ref{cor_MTL_loss} can be found in Appendix \ref{app_iso_cov}.
As a remark, %in equation \eqref{cor_MTL_error}, the predication loss $L(\hat{\beta}_2^{\MTL}) $ was obtained using the global minimizer $(\hat A,\hat B)$. B
by combining the bias and variance limits, we can also obtain a bias-variance tradeoff for any local minimizer of $f(A, B)$.
The proof is similar to Corollary \ref{cor_MTL_loss}, so we omit the details.

@@

\begin{example}[Covariate shift]\label{ex_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example focuses on how varying covariate shifts impacts the \textit{variance} limit in equation \eqref{lem_cov_shift_eq}. For large enough $p$,
%in the left hand side of equation \eqref{lem_cov_shift_eq}:
\begin{align*}
	\bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} &\rightarrow \frac{1}{n_1 + n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}} \\
	&= \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}.
\end{align*}
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Hence the variance limit is determined by the spectrum of $M$. % To be clear, for this example we assume that the bias is $0$.
To illustrate the above result, suppose that half of $M$'s singular values are equal to $\lambda > 1$ and the other half are equal to $\lambda^{-1}$.
In particular, when $\lambda = 1$, there is no covariate shift.
As $\lambda$ increases, the severity of covariate shift increases.
We observe the following dichotomy.
%\begin{itemize}

@@

We explain why the dichotomy happens. The variance limit in this example is equal to
$\frac{p}{2(n_1 + n_2)} f(\lambda)$, where
\[ f(\lambda) = {(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}. \]
Using the fact that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2}$, we can verify
%\begin{align*}
%	f(\lambda) - f(1) &= \left(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&+ \left(\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&- \frac{2(n_1 + n_2)}{n_1 + n_2 - p} \\
%	&= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
%\end{align*}
\begin{align*}
	f(\lambda) - f(1) &= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
\end{align*}
where $g(\lambda, a_1) \ge 0$.
%and can be derived from algebraic calculations (details omitted).
We claim that $a_1 \ge \frac{n_1 + n_2-p}{2(n_1 + n_2 )}$ if and only if $n_1 \ge n_2$, which explains the dichonomy. In fact, if $a_1>a_2$, then equation \eqref{eq_a12extra000} gives that $a_1> \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$, and equation \eqref{eq_a12extra} gives that
\begin{align*}
 \frac{n_1}{n_1 + n_2} &> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) >\frac{1}{2}.
\end{align*}
%\begin{align*}
% \frac{n_1}{n_1 + n_2} &=a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} \\
%	&> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) =\frac{1}{2}.
%\end{align*}
This implies $n_1>n_2$. The other direction follows from similar arguments. % Similarly, if $a_1<a_2$, equations  \eqref{eq_a12extra000} and  \eqref{eq_a12extra} give that $a_1 < \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$ and $n_1<n_2$.
%Thus, we conclude that $f(\lambda) \ge f(1)$ if and only if $n_1 \ge n_2$.

@@

\textbf{Proof overview of Theorem \ref{thm_main_RMT}.}
For the rest of this section, we present an overview of the proof of Theorem \ref{thm_main_RMT}.
The central quantity of interest is the inverse of the sum of two sample covariance matrices.
We note that the variance equation $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}]$ is equal to $(n_1 + n_2)^{-1} \bigtr{W^{-1}}$, where $W$ is
\begin{align}\label{eigen2extra}
	\frac{\Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V}{n_1 + n_2}.
\end{align}
Here $U\Lambda V^\top$ is defined as the SVD of $M$.
This formulation is helpful because we know that $(Z^{(1)})^{\top} Z^{(1)}$ and $(Z^{(2)})^{\top} Z^{(2)}$ are both sample covariance matrices with isotropic population covariance, and $U, V$ are both orthonormal matrices.
For example, if $Z^{(1)},Z^{(2)}$ are both Gaussian random matrices, by rotational invariance, $Z^{(1)} U, Z^{(2)}V$ are still Gaussian random matrices.

@@

\begin{figure*}[!t]
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/same_covariates.eps}
		\caption{Example \ref{ex_same_cov}}
		\label{fig_same_cov}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/sample_ratio_several_d.eps}
		\caption{Example \ref{ex_sample_ratio}}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/covariate_shift.eps}
		\caption{Example \ref{ex_covshift}}
		\label{fig_covariate}
	\end{subfigure}
	\caption{%Three takeaways of our theory in Section \ref{sec_insight}.
	Our estimated losses (solid line) match the empirical losses (dotted line) accurately under various settings in dimension $p = 200$.
	\textbf{Left.} Validating Example \ref{ex_same_cov} for ten tasks: the noise variance $\sigma^2$ is $1/4$.
	\textbf{Middle.} Validating Example \ref{ex_sample_ratio} for two tasks: we discover an interesting phenomena by fixing task two's sample size and increasing task one's sample size.
	Moreover, our result accurately predicts the critical point (marked in circle) of the loss curve.
%	Depending on how large the distance $d^2$ is, task two's prediction loss decreases initially before increasing again, or decreases monotonically.
	\textbf{Right.} We show how different levels of covariate shift affect hard parameter sharing when there is no bias.
	Having covariate shift increases task two's prediction loss when task two's sample size is smaller than task one. Otherwise, having covariate shift (surprisingly) decreases task two's prediction loss.}
	\label{fig_model_shift_phasetrans}
\end{figure*}

### 2020-10-15 arxiv/experiments.tex
polish

\textbf{Sample efficiency.}
First, we study the setting of Example \ref{ex_same_cov}.
Figure \ref{fig_same_cov} shows the average prediction loss as we increase the number of samples per-task   from $400$ to $2000$.
In all the parameter settings, our results estimate the empirical losses accurately.
We also observe a trend that the average prediction loss increases as we increase $d$ from $0.1$ to $0.2$.
Our work explains the differences between these two settings since $d^2 = 0.1^2$ is always smaller than $\frac{\sigma^2 p}{n - p}$, but $d^2 = 0.2^2$ is not.
%Second, when $d = 0.1$, we have that $d^2 \le \frac{\sigma^2 p}{n - p}$ for all values of $n$, hence the average prediction loss of hard parameter sharing is always lower than STL.
Indeed, we observe a crossover point between hard parameter sharing and STL.
Finally, for $d = 0.2$, looking horizontally, we find that HPS requires fewer samples per-task than STL to achieve the same loss level. %\FY{I do not quite understand this sentence about "3x fewer", because how much data needed depends on $d$ and the prediction loss level we are looking at. For example, at the cross point, this ratio is 1. }

\textbf{Sample ratio.}
Second, we show how varying sample ratio impacts hard parameter sharing in the setting of Example \ref{ex_sample_ratio}.
Figure \ref{fig_size} shows task two's prediction loss as we increase the sample ratio $n_1 / n_2$ from $1/10$ to $7/10$.
%Again, our estimates are accurate compared to the empirical losses.
We consider a regime where task two consists of $80,000$ samples, and task one's sample size varies from $8,000$ to $56,000$. 
The task-specific variance is $d^2 = 0.08$, the noise variance is $\sigma^2 = 0.3^2$, and the shared signal variance is $1$. We observe that as we increase the sample ratio, task two's prediction loss decreases initially but later will increase when the sample ratio is above a certain level.
On the other hand, when $d^2 = 0.045$, task two's prediction loss decreases monotonically.
Our result from Corollary \ref{cor_MTL_loss} explains this trend. \HZ{check}

\textbf{Covariate shift.}
Finally, we show how varying covariate shift impacts hard parameter sharing in the setting of Example \ref{ex_covshift}. 
Figure \ref{fig_covariate} shows task two's prediction loss as we increase task one's sample size.
Recall that $\lambda$ measures the severity of covariate shifts---a larger $\lambda$ means a larger covariate shift.
We indeed observe the dichotomy in Example \ref{ex_covshift} at $n_1 = 800$.
The sample size $n_2$ is $800$ and the noise variance $\sigma^2$ is $1/4$.
%Covariate shift: We set $\kappa = 1$ and $d = 0$.
%We set $\rho_2 = 4$ and vary $\rho_1$ from $5$ to $25$ for sample sizes.
%We use the scale parameter $\lambda = 1$ for the curve without covariate shift and $\lambda = 2$ for the curve with covariate shift (cf. Section \ref{sec_covshift}).

>>>>

\textbf{Sample efficiency.}
First, we validate the result of Example \ref{ex_same_cov}.
Figure \ref{fig_same_cov} shows the average prediction loss as we increase the number of samples per-task   from $400$ to $2000$.
In all the parameter settings, our results estimate the empirical losses accurately.
We also observe a trend that the average prediction loss increases as we increase distance $d$ from $0.1$ to $0.2$.
Our work explains the differences between these two settings since $d^2 = 0.1^2$ is always smaller than $\frac{\sigma^2 p}{n - p}$, but $d^2 = 0.2^2$ is not.
%Second, when $d = 0.1$, we have that $d^2 \le \frac{\sigma^2 p}{n - p}$ for all values of $n$, hence the average prediction loss of hard parameter sharing is always lower than STL.
Indeed, we observe a crossover point between hard parameter sharing and STL.
Finally, for $d = 0.2$, looking horizontally, we find that HPS requires fewer samples per-task than STL to achieve the same loss level. %\FY{I do not quite understand this sentence about "3x fewer", because how much data needed depends on $d$ and the prediction loss level we are looking at. For example, at the cross point, this ratio is 1. }

\textbf{Sample size ratio.}
Second, we validate the result of Example \ref{ex_sample_ratio}.
Figure \ref{fig_size} shows task two's prediction loss as we increase the sample ratio $n_1 / n_2$ from $1/10$ to $7/10$.
%Again, our estimates are accurate compared to the empirical losses.
We consider a regime where task two consists of $80,000$ samples, and task one's sample size varies from $8,000$ to $56,000$. 
The task-specific variance is $d^2 = 0.08$, the noise variance is $\sigma^2 = 4^2$, and the shared signal variance is $1$. We observe that as we increase the sample ratio, task two's prediction loss decreases initially but later will increase when the sample ratio is above a certain level.
On the other hand, when $d^2 = 0.045$, task two's prediction loss decreases faster.
Intuitively, this is because bias increases less for smaller $d^2$.
%Our result from Corollary \ref{cor_MTL_loss} explains this trend.

\textbf{Covariate shift.}
Finally, we validate the result of Example \ref{ex_covshift}.
Figure \ref{fig_covariate} shows task two's prediction loss as we increase task one's sample size.
Recall that $\lambda$ measures the severity of covariate shifts---a larger $\lambda$ means a larger covariate shift.
We indeed observe the dichotomy in Example \ref{ex_covshift} at $n_1 = 800$.
The sample size $n_2$ is $800$ and the noise variance $\sigma^2$ is $1/4$.
%Covariate shift: We set $\kappa = 1$ and $d = 0$.
%We set $\rho_2 = 4$ and vary $\rho_1$ from $5$ to $25$ for sample sizes.
%We use the scale parameter $\lambda = 1$ for the curve without covariate shift and $\lambda = 2$ for the curve with covariate shift (cf. Section \ref{sec_covshift}).

### 2020-10-15 arxiv/intro.tex
polish

Hard parameter sharing (HPS) for multi-task learning is widely used in empirical research and goes back to the seminal work of \citet{C97}.
Recent work has revived interest in this approach because it improves performance and reduces the cost of collecting labeled data \cite{R17}.
It is generally applied by sharing the feature layers between all tasks while keeping an output layer for every task.
Often, hard parameter sharing offers two critical advantages if successfully applied.
(i) It reduces model parameters since all tasks use the same feature space.
(ii) It reduces the amount of labeled data needed from each task by augmenting the entire training dataset.

Hard parameter sharing works as an inductive transfer mechnism and a regularizer that reduces overfitting, both of which have great intuitive appeal \cite{R17}.
For example, by restricting the shared space's size, HPS encourages information sharing among multiple tasks \cite{KD12}.
Another source of inductive bias comes from the tasks and depends on datasets' properties such as sample sizes and task covariances \cite{WZR20}.
However, how these dataset properties impact HPS has not been well established.
%It becomes increasingly important to understand HPS' formal generalization properties.
Part of the challenge may be that HPS' generalization performance depends intricately on the sample size ratios and covariate shifts between tasks, and is not amenable to standard concentration results.
Previous results based on Rademacher complexity or VC dimensions have considered when all tasks' sample sizes are equal to logarithmic factors of the feature dimension \cite{B00,MPR16}, and when all tasks' sample sizes increase simultaneously \cite{AZ05,M06}.
%For, the generalization error scales down as the sample sizes of all tasks increase, when applied to the multi-task setting \cite{B00,AZ05,M06,MPR16,WZR20}.

>>>>

Hard parameter sharing (HPS) for multi-task learning is widely used in empirical research and goes back to the seminal work of \citet{C97}.
Recent work has revived interest in this approach because it improves performance and reduces the cost of collecting labeled data \cite{R17}.
It is generally applied by sharing the feature layers between all tasks while keeping an output layer for every task.
Often, hard parameter sharing offers two critical advantages if successfully applied.
First, it reduces model parameters since all tasks use the same feature space.
Second, it reduces the amount of labeled data needed from each task by augmenting the entire training dataset.

Hard parameter sharing works as an inductive transfer mechnism and a regularizer that reduces overfitting, both of which have great intuitive appeal \cite{R17}.
For example, by restricting the shared space's size, HPS encourages information sharing among multiple tasks \cite{KD12}.
Another source of inductive bias comes from the tasks and depends on datasets' properties such as sample sizes and task covariances \cite{WZR20}.
However, how these dataset properties impact HPS has not been well established.
%It becomes increasingly important to understand HPS' formal generalization properties.
Part of the challenge may be that HPS' generalization performance depends intricately on the sample size ratios and covariate shifts between tasks, and is not amenable to standard concentration results.
Previous results based on Rademacher complexity or VC dimensions have considered cases where all tasks' sample sizes are equal to logarithmic factors of the feature dimension \cite{B00,MPR16}, and when all tasks' sample sizes increase simultaneously \cite{AZ05,M06}.
%For, the generalization error scales down as the sample sizes of all tasks increase, when applied to the multi-task setting \cite{B00,AZ05,M06,MPR16,WZR20}.
